### 2022-03-14 response_reviewers.tex

\textbf{Reviewer \#1:}
I greatly enjoyed reading the paper "Ten simple rules to host an inclusive conference", and think it provides an easy to follow set of rules that organizers can use to ensure an inclusive conference. I found many tips for inclusiveness in the article to be especially helpful, such as multilingual accommodations and ensuring screen-reader accessibility.

I just have a few aspects I wasn't 100\% clear on.

(1) In Rule 4, the manuscript states "Define measurable indicators for these goals e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions", but how do organizers choose the correct demographics to monitor and the expected background rates? Should it match global, the local, or field-specific rates? Or should it be remediative, where it specifically selects for communities historically ignored in the field beyond the regional population rate?

\begin{Reply}
   We are going to refer to line numbers in the following way: In the plos\_latex\_template.tex document, add a line number label where you want to edit like this: \\ linelabel\{rev1-com1-a\}
   To refer to these lines in this document, add a reference here, like this: Lines   \ref{rev1-com1-a} to \ref{rev1-com1-b}

   %ast: We do not have a correct answer for this, unfortunately, and we believe there is a reflexion work that every team has to do in their context. We had extensive discussions about the indicators and our conclusion is that the "correct" demographic to monitor will vary for every conference, field, and local and regional context. 
   % Depending on the field, greater achievements can be made, and in some areas, this exercise will be more difficult and yet extremely important even if the apparent advancements are less evident. 
   % Our team is inclined to propose what the reviewer calls "remediative measures", or affirnative actions, and be explicitly more than inclusive, but we understand that not every team can set such targets. And even with bold intentions, there is no guarantee that the final result is going to be as expected. 
   % A support from the community may be a great way to strenghen bold goals. However, in some communities the common sense shared by members may represent a wish to preserve the status quo, so not always a broad support from the community can be expected -but this may indicate also a need to broaden the definition of the community.
   % We would like to encourage the readers of any field and regional context to make this exercise independently of the status of the field and the initial impression that these efforts won't be successful or sufficient. 
   % to edit: We have made this explicit in the text. We also included the fact that the community 1. should be defined and 2. can be consulted before the conference to make sure that this construction is collective and participative.

%Marce: I agree with reviewer 1 that this should be remediative, quoting: "where it specifically selects for communities historically ignored in the field beyond the regional population rate". To respond to the same reviewer, I think we can mention how  UseR presented the demographics of the attendees/participants, and how this was shared with the community as a way to be held accountable.
%Marce: About rule 10: both reviewers 1 and 3 suggested to discuss further how to implement these rules. I think rule 10 is already discussing that in a clear manner.
\end{Reply}

@@

(2) Adding on to the first point, how are the indicators measured? Should the conference organizers quantify the demographics of the attendees/participants? If so, should these details be published to the participants after the conference? How do organizers hold themselves accountable? And to whom are they held accountable to; the organizing committee themselves or the greater community?

\begin{Reply}
   % Yani: I think we can mention and link to the info board and to the blog post post conference analyzing this numbers like the Latin American one and the accessibility one. The surveys are also another source of data for this accountability. I think this is related with the other review asking about governance. 
   %ast: The tenth rule is directly connected to the first, because these indicators can be measured and evaluated via surveys, direct impressions from the atendance, and exercises such as focus groups. Of all these options, surveys are the most frequently used but we wanted to highlight that there are other options. This link to rule 10 was not explicit in the text (edit!)
   % About publishing the details after the conference, the privacy and anonymity of the participants should be respected, so some kind of binning/grouping of the data and care with categories that have a low number of respondents should be taken. %make sure this in the manuscript! we had this in previous versions.
   % Accountability of the team should be posible with or without public indicators. Open channels for communication and open fields and the possibility to make anonymous comments should be provided. Especially if there is a claim for diversity and inclusion, the general community should be able to give feedback about the success of these efforts. 
   %to edit: cite the tenth rule in the first, telling there will be accountability. but also say that the indicators should not become a goal themselves.

\end{Reply}
(3) In Rule 4, it states "Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe;" Does this pertain to invited speakers only, or speakers selected from submitted abstracts? If this pertains to scored abstracts, how does one ensure that this is done in a transparent and reproducible way? Or maybe I'm not fully grasping what curate means in this context?

\begin{Reply}
   %ast: Although curation is easier and definitely encouraged for invited speakers, we would like to reinforce that even transparent scored processes can be influenced by implicit bias, once the reviewers start evaluating there might be "shifting criteria" and other considerations entering mid-process.
   %We have added some extra explanation about this curation, that include at least: 
   %1. determining clear criteria before seeing the list of submissions, 2. examining the final list of selected work for imbalances in gender, geographical representation and other visible aspects of diversity. 3. check if close competitors (similar quality work) were excluded for some apparent bias.
   % This process is hard because there are many aspects of diversity that will be unreachable for reviewers and are prone to perceived cues rather than real information (so not every dicersity dimension can be considered here). The invitation is to examine similar quality work and try to understand why some work was included rather than other and to consider whether bias could have influenced the outcome.
   %criteria should be clear before the abstract evaluation. Some adaptation of this can be made: https://honors.agu.org/files/2014/11/AWARDS-SUGGESTED-BEST-PRACTICES-MAY2011.pdf
\end{Reply}

(4) In Rule 5, it states "The physical venue would have to include multiple ways to connect with online folks such as cameras and microphones to allow them to follow who is speaking at the in-person space, and tablets to help in-person attendees interact with remote participants." How do the tablets help in-person attendees interact with remote participants? Especially participants that may be in a different time-zone. Or are the tablets used to help the online activities mentioned in the next paragraph?

\begin{Reply}
   %ast: Although one can assume that an ever growing proportion of people have their own handheld device to interact online with the remote participants, we cannot assume that everyone does, or that the models are compatible with the options of the conference. Some devices (and support for those and the participants') should be provided by the organization. With everyone having a device, the idea is to unify the networking activities and Q & A sessions to be online, this more integrated with the remote component.
\end{Reply}

\begin{Reply}
%ast: "and" should be "or". Yes, definitely, rather than two abstracts, one in English and the other in a second language, the best way would be to be able to have abstracts in languages different than English without the need for a duplication. That depends on the conference capabilities but we would recommend both submission and presentation be open to other languages. The idea for a previous request/screening about the languages that will be needed is good. Another option is to have a language editing help for English, via the communities or via the organizing team, that helps ensure that the submitted abstracts are conveying the meaning the authors are trying to give. This could include also, abstracts written in English by students and early career researchers. 
\end{Reply}

\begin{Reply}
%ast: We thank the reviewer's feedback. We have added some pre-conference participation of the community, also inspired by examples we got from colleagues at pycon Germany. The accountability is very hard to define because no formal structure requires that in the present, it has to be the team's will to be held accountable to their participant community. Clarity in the communication strategy and open spaces for feedback are key here.
\end{Reply}

\textbf{Reviewer \#2:}
L78-L80 As a female scientist I would ask myself "where are we" (because we still have parts of the World where female scientist are rare on leading positions), but I consider science apolitical in every segment, and I certainly suggest that it concentrate here on the scientific youth, the segment of research in industry, etc., and not on the above.
The rest of the paper it should certainly be praised because it enters into the essence of the exchange of knowledge today.

@@

\textbf{Reviewer \#3:} 
This is a timely, well-written commentary on organising an inclusive conference and should act as a reference compendium for researchers, practitioners as well as conference organisers.

I have some comments for authors to consider, and possibly expand upon.

1. As much as I enjoyed reading the ten rules, I missed finding mention of how this is to be implemented and governed. Therefore, it is poignant that authors should brief discuss having an inclusion oversight or governance committee that tracks these metrics and their implementation.

\begin{Reply}
%ast: my first hunch in general: the metrics part is one of the hardest part to implement and reviewer1 talked about it, but it is also one of the "goals" that stick with some readers and we were trying to avoid putting them at the center of decision-making because they can become a end in itself and not a way to understand if your efforts were successful
%The tenth rule should have accounted for this but we may need to be more specific in the messages of not letting yourself be prey of the indicators. 
%Team accountability and governance were part of rule 10 so maybe we could strengthen this.
\end{Reply}

@@

2. Secondly, I also missed detailed methodology on how these ten rules were developed or consensus were reached at? Were stakeholders from diverse backgrounds consulted? Was the consultation process iterative? Were the recommendations thus made by those from diverse backgrounds incorporated? How were the final draft reached at - through consensus or voting or a process to that effect? This is important as for such a set of indications to be indicative, it is important that opinions from all backgrounds are taken into account and co-developed.

\begin{Reply}
   %Yani: I think this is not on the scope of the paper, but we can give an explanation of the process, which I think was fantastic and inclusive and have a diverse background on the construction of the rules. 
   %ast: i agree, this may break the format for a ten simple rules piece so maybe a very short mention in the introduction or as SupMat?
\end{Reply}

@@

3. In Rule 1, though gender or ethnic diversity has been highlighted, geographical diversity needs to be discussed in detail. There is significant "north-south" divide in scientific research - with several of institutions as well as researchers from LMICs or global south not being given appropriate representation on conference committees as well as on speakers diversity.

4. In Rule 2, safe and inclusive environments, I find little attention to "racist" aberrations or incidents - and how they should be mitigated and when reported be appropriately addressed with fairness and justice.
Several speakers from ethnically diverse backgrounds face racism, exclusion or apathy at such conferences. This need to be discussed.

%ast: All specific examples of offenses should be included in the code of conduct construction process, and that's why we didn't make a list here, but maybe we can make explicit mentions of what type of offenses can/have to be in the code of conduct. %suggestion for edits: add this, talk a little bit more about the content of the CoC text.

@@

%ast: All specific examples of offenses should be included in the code of conduct construction process, and that's why we didn't make a list here, but maybe we can make explicit mentions of what type of offenses can/have to be in the code of conduct. %suggestion for edits: add this, talk a little bit more about the content of the CoC text.

\end{Reply}

5. In Rule 4, via a vis bias, bias against people with disability and ethnic minorities should also be highlighted.

\begin{Reply}
%ast: We have tried to make a distinction between visible and invisible aspects of human diversity, and we have to respect the right of every individual to occupy the space and assert/disclaim who they are and the right to remain silent about some aspects of their lives. Rule 2 is about being proactive regarding all dimensions of diversity, and Rule 4 focuses more on the visible aspects of diversity, especifically those that can be grasped via written communication.
% We mention this because counteracting bias actions during the selection of speakers and submitted work, we--unfortunately--only have *perceived* cues of these dimensions. We can cite Laura's piece on video abstracts being exclusionary and be more explicit about the fact that perceived cues and implicit and explicit racism go hand in hand.
\end{Reply}

@@

6. In Rule 8, inclusive communication strategy - it is important that the speakers from diverse backgrounds, especially from ethnic minorities etc are highlighted or given prominence in social media strategy or communication. Often, these speakers face relatively less focus. This may be discussed.

\begin{Reply}
   %Yani: I think all keynote deserve the _same_ communication, if we give prominence to minoritized group could be perceived as tokenism. So I would take the suggestion on NOT give more prominence to one group of keynote and make sure that all of them are promote in the same spaces and in the same way.  We can mention here that communities partners can help with the highlight of the minoritazed groups (tricky).
   %ast: I agree with Yani, and we had it partially covered in the rule 7, "Promote these sessions among all conference attendees and announce if captions will be available; non-English sessions should be given the same relevance as the rest of the program." but we can add something here. 
\end{Reply}

### 2022-03-17 plos_latex_template.tex
Comments by @jupanago and @rociojoo, start editing by @andreasancheztapia

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects included in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it's good to have them in mind from the beginning.
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably.
Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

@@

The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. 
Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.

@@

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted. 
Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing all of the submission information to
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
Define clear evaluation criteria and share them with the reviewers and authors.
\hl{Check the final lists of speakers and do not hesitate to curate them to correct any imbalances you observe}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.

>>>>

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects included in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and great advice on `low-hanging fruits'), some others need careful planning in advance, so it's good to have them in mind from the beginning.
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM\hl{, women in science}\linelabel{wis}), and sessions aimed to welcome first-timers to the conference.

@@

% Rocío: Idea of this paragraph: tips to make a diverse and inclusive team work + paying them
If you have already started assembling an organizing team, check for gaps in its composition and do your best to fill them. 
The regional and local communities, groups, or associations in your field are good sources to tap into. 
When it comes to team leadership, do not expect self-nomination and voting to work as infallible mechanisms to fill these positions equitably. \linelabel{team}
Instead, nominate folks directly and offer them leadership positions in the organization team, especially those positions that would be typically occupied by people from privileged groups.
Bear in mind that some people may lack the institutional or financial support to put time and effort into the organization tasks and do not have the luxury to commit for free; consider reallocating resources to give them a stipend for their work (see \textbf{Rule 9}).
In any case, try to find meaningful ways to acknowledge everybody's work. Use the conference spaces (e.g. webpage, opening/closing ceremony at the conference, social media) to give proper credit to the people who are making the event possible.
And most importantly, take care of your team and their well-being! Check on them regularly and make sure that everyone is comfortable. This represents a great amount of work and you will need to support each other in the long run.

@@

The conference program may lack diversity in the invited roles––such as keynote speakers, program/scientific committee, session chairs, panelists––, the accepted abstracts, and even the scope of the topics covered. 
It is important to counteract biases in all of these aspects.
When choosing or inviting people for visible and valued roles in the conference, it is likely that there will not be much diversity in the first pool of names \cite{dwyerNoticeWhoScience2021,swartzScienceValueDiversity2019,wongBuildDiversityScience2020,dignazioUnicornsJanitorsNinjas2020}. \linelabel{invited}
Rather than deterring you, this should encourage you to go beyond your current networks to look for, reach out to, invite, encourage, and onboard great people that are not routinely in the spotlight, making sure that these roles do not reinforce existing privileges.

@@

The abstract review process can be subjected to implicit bias \cite{ross_everyday_2020} when folks in the selection committee unconsciously assign a positive or negative value to the names and affiliations of authors—and their perceived origin, ethnicity, gender, or native language—, affecting the chances to get their work accepted. 
Selection committees need to be reminded of such biases in the evaluation process to try to avoid judging more scrupulously the work of people perceived as part of minoritized groups, or being kinder when reviewing the abstracts from those perceived as part of privileged groups and considering their work as more relevant \cite{swartzScienceValueDiversity2019, andersson_implicit_2019}.
Evaluate the plausibility of anonymizing \hl{as much of the submission information as possible} \linelabel{anonymous}, and
define the most appropriate strategy for reviewing (e.g. double blind, double open, or single blind reviews; see \cite{numfocus_discover_2021}).
Define clear evaluation criteria and share them with the reviewers and authors \hl{before they read any of the submissions} \linelabel{criteria}.
\hl{Examine the final list of selected abstracts to verify if similar quality work may have been judged differently due to some apparent bias} \linelabel{not-curation}; even well-meaning and carefully crafted processes can be prone to reproduce the biases present in academia and society.
